\documentclass[12pt, oneside, a4paper]{book}

\usepackage[pdftex]{graphicx}
\usepackage{tikz-cd}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{pgfplots}
\usepackage{xcolor}
\usepackage[shortlabels]{enumitem}
\usepackage[mathscr]{euscript}
\pgfplotsset{compat = newest}
\usepackage{float}
\usepackage{microtype}
\usepackage[mode=text]{siunitx}

\usepackage{vuwproposal} % sets up some local things, mostly the front page

\usepackage{palatino} % sets palatino as the default font

\usepackage{url} % for typesetting urls

\usepackage{booktabs} % for table creation



%\renewcommand{\baselinestretch}{1.00}


\begin{document} 
	
	
\frontmatter
% Book style knows about front matter
% Report style doesn't so you need to set roman numbering etc yourself :-(

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	
\title{Ordinal Pattern Features: Statistical Properties and Their Applications in Time Series Clustering}
%%% ACF Correct the title
%%% ACF Suggestion: Statistical Properties of Features from Ordinal Patterns and Their Application to Time Series Clustering

\author{Rasika Dilhani}
\supervisor{Alejandro C.\ Frery}
	
\subject{Data Science}
\abstract{Time series analysis plays a vital role in understanding the underlying dynamics of complex systems across various domains such as engineering, economics, and the physical sciences. Traditional approaches, namely time-domain and frequency-domain methods often rely on strong assumptions such as stationarity, large sample sizes, or normality, which are frequently violated in real-world data. As a robust alternative, Bandt and Pompe~\cite{PhysRevLett.88.174102} introduced a non-parametric method based on ordinal pattern symbolization and information theoretic descriptors. 
This approach transforms local segments of the time series into rank-based symbols (the ordinal patterns), 
constructs a histogram of ordinal patterns, and 
computes Shannon entropy, offering resistance to noise and model independence. 
%%% ACF 1- symbolisation, 2- histogram, 3- descriptors (H and C)	
In addition to the foundational work by Bandt and Pompe, we refer to the contributions of López-Ruiz et al. \cite{lopez1995statistical}, who introduced the concept of statistical complexity; Lamberti et al. \cite{lamberti2004intensive}, who implemented López-Ruiz’s idea using the Euclidean distance; Rosso et al. \cite{EEGAnalysisUsingWaveletBasedInformationTools}, who proposed the joint representation of entropy and complexity in the entropy-complexity plane; and Martin et al. \cite{Martin2006}, who established the theoretical boundaries of the generalized statistical complexity measure. 
This proposal investigates the use of various entropy measures, such as Shannon, Tsallis, and Rényi entropy, as well as the Fisher Information Measure, statistical complexity, and associated confidence intervals for time series analysis clustering. 
%The focus is on converting time series data into symbols by analyzing their ordinal patterns, constructing pattern histograms, estimating descriptors such as entropy and complexity, calculating confidence intervals for these descriptors, and finally analyzing the entropy-complexity plane to draw conclusions.
%%% ACF What do we do with respect to histogram construction?
%The final conclusion is based on the analysis of real-world data using the investigations mentioned above.
} 


% Books don't normally have abstracts, and this is a bit of a hack
	
% Uncomment the appropriate degree
\phd
%\mscthesisonly
%\mscwithhonours
%\mscbothparts
% \otherdegree{DEGREE OR DIPLOMA NAME}
	
	
	
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%New commands:
	
\newcommand{\Sha}{\rotatebox[origin=c]{180}{$\Pi\kern-0.361em\Pi$}}
\newcommand{\ot}{\otimes}
\newtheorem{definition}{Definition}
\newtheorem{theo}{Theorem}
\newtheorem{contoh}{Example}
\newtheorem{cor}{Corollary}
\newtheorem{prop}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{remark}{Remark}
\newtheorem{conjecture}{Conjecture}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	
\maketitle
	
%\include{thesis-acknowledge}
	
\tableofcontents
	
	
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	
% book style knows about mainmatter
% if you are using report style you will have to rest page numbering etc.
\mainmatter
	
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	
% individual chapters included here
	
\include{Chapter1}
\include{Chapter2}
\include{Chapter3}
\include{Chapter4}
\include{Chapter5}
	
	
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	
% and of course book style knows about backmatter
% \backmatter caused problems with appendices :-(
% and of course report style doesn't
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	
	
%\bibliographystyle{ieeetr}
\bibliographystyle{acm}
\bibliography{BearingFaultDiagnosis}
	
	
\end{document}
